Well-quasi-orderings on word languages

### 2025-10-16 src/conclusion.tex
fix: knowledges + refs + storytelling

However, there are \kl{downwards closed} languages that are
\kl{well-quasi-ordered} for the \kl{infix relation} but have an \kl{exponential
factor complexity}: the $(5, 3)$-Toeplitz word is \kl{uniformly recurrent}
\cite[p. 499]{CAKA97}, but has an exponential factor complexity \cite[Theorem
5]{CAKA97}. This shows that our computational models somehow fail to capture
vast classes of \kl{well-quasi-ordered} languages with a high \kl{factor
complexity}. It would be interesting to understand which new proof techniques
would be required to obtain decidability for these languages.

>>>>

However, there are \kl{downwards closed} languages that are
\kl{well-quasi-ordered} for the \kl{infix relation} but have an exponential
\kl{factor complexity}: the $(5, 3)$-Toeplitz word is \kl{uniformly recurrent}
\cite[p. 499]{CAKA97}, and has exponential factor complexity \cite[Theorem
5]{CAKA97}. This shows that our computational models somehow fail to capture
vast classes of \kl{well-quasi-ordered} languages with a high \kl{factor
complexity}. It would be interesting to understand which new proof techniques
would be required to obtain decidability for these languages.

### 2025-10-16 src/introduction.tex
fix: knowledges + refs + storytelling

\AP Let us argue that these decision procedures fit a larger picture in the
research area of well-quasi-orderings.
Indeed, there has been recent breakthroughs in deciding whether a given order
is a \kl{well-quasi-order}, for instance in the context of the verification of
infinite state transition systems~\cite{DBLP:conf/fsttcs/FinkelG19} or in the
context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
In the graph theory community, recent works have studied classes of graphs 
that are \kl{well-quasi-ordered} by the \kl{induced subgraph relation}
using similar language theoretic techniques~\cite{DRT10,lopez25,ALM17}.
Furthermore, a previous work by Kuske shows that any
\emph{reasonable}\footnote{ This will be made precise in
\cref{infix-embedding:thm}. } partially ordered set $(X, \leq)$ can
be embedded into $\set{a,b}^*$ with the \kl{infix relation}~\cite[Lemma
5.1]{DBLP:journals/ita/Kuske06}. Phrased differently, one can encode a large
class of partially ordered sets as subsets of $\set{a,b}^*$. As a consequence,
the following decision problem provides a reasonable abstract framework for
deciding whether a given partially ordered set is \kl{well-quasi-ordered}:
given a language $L \subseteq \Sigma^*$, decide whether $L$ is
\kl{well-quasi-ordered} by the \kl{infix relation}.

>>>>

\AP Let us argue that these decision procedures fit a larger picture in the
research area of well-quasi-orderings.
Indeed, there has been recent breakthroughs in deciding whether a given order
is a \kl{well-quasi-order}, for instance in the context of the verification of
infinite state transition systems~\cite{DBLP:conf/fsttcs/FinkelG19} or in the
context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
In the graph theory community, recent works have studied classes of graphs 
that are \kl{well-quasi-ordered} by the induced subgraph relation
using similar language theoretic techniques~\cite{DRT10,lopez25,ALM17}.
Furthermore, a previous work by Kuske shows that any
\emph{reasonable}\footnote{ This will be made precise in
\cref{infix-embedding:thm}. } partially ordered set $(X, \leq)$ can
be embedded into $\set{a,b}^*$ with the \kl{infix relation}~\cite[Lemma
5.1]{DBLP:journals/ita/Kuske06}. Phrased differently, one can encode a large
class of partially ordered sets as subsets of $\set{a,b}^*$. As a consequence,
the following decision problem provides a reasonable abstract framework for
deciding whether a given partially ordered set is \kl{well-quasi-ordered}:
given a language $L \subseteq \Sigma^*$, decide whether $L$ is
\kl{well-quasi-ordered} by the \kl{infix relation}.
